---
title: "Reproducible chat instructions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducible chat instructions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = FALSE,
  message = FALSE,
  warning = FALSE
)
```

This vignette shows how to:

- discover which instruction modules are shipped with `{reproducibleai}`,
- use recommended compositions (“recipes”),
- write selected instruction files into a local folder (e.g. `dev/instructions/`) so future chat sessions are reproducible,
- use the `CHAT_INSTRUCTIONS.md` entrypoint to ensure chats follow the intended instruction modules.

## Why instruction modules?

A reproducible chat session needs two things:

1. **Stable guidance** (what rules the assistant should follow), and
2. **A local record** of exactly which guidance was used.

`use_instructions()` supports (2) by copying the chosen instruction modules into a known folder.

## List available modules

Use `instructions_available()` to see what modules are installed with the package:

```{r}
library(reproducibleai)

instructions_available()
```

If you want the resolved file paths:

```{r}
instructions_available(include_path = TRUE)
```

## Use recommended compositions (recipes)

`instructions_recipes()` provides opinionated, maintained “known good” compositions.

```{r}
recipes <- instructions_recipes()
names(recipes)

recipes$base
recipes$r_package
recipes$quarto_user_manual
```

## Write instructions into a project folder

In a real project, you’ll usually write to `dev/instructions/`:

```r
use_instructions(c("chat-manual", "goals", "r-package"))
```

For this vignette, we write into a temporary directory so nothing in your working directory is modified:

```{r}
tmp <- tempdir()
dest <- file.path(tmp, "dev", "instructions")

paths <- use_instructions(
  spec = c("chat-manual", "goals", "quarto-book", "user-manual"),
  dest_dir = dest,
  quiet = TRUE
)

paths
list.files(dest)
```

You can inspect the written files to confirm what guidance was used:

```{r}
cat(readLines(file.path(dest, "chat-manual.md"), warn = FALSE)[1:10], sep = "\n")
```

## The `CHAT_INSTRUCTIONS.md` entrypoint

Once your selected modules have been written to `dev/instructions/`, a chat session can start from a single entrypoint file:

- `dev/instructions/CHAT_INSTRUCTIONS.md`

The entrypoint file documents:
- which instruction modules apply to this repository, and
- the order they should be read/applied.

This “entrypoint” pattern makes it easy to start a new chat session reproducibly: you refer to **one file**, and that file declares the selected modules.

## A practical workflow

A typical “start of day” workflow might look like:

1. Choose a composition:
   - for Quarto technical manuals: `c("chat-manual", "goals", "quarto-book", "user-manual")`
   - for Shiny golem apps: `c("chat-manual", "goals", "r-package", "shiny-golem")`
2. Write the modules into the project:
   - `use_instructions(spec)`
3. Keep the resulting files under version control:
   - commit `dev/instructions/*.md` so reviewers can see exactly what guidance governed the chat session.
4. Start future chat sessions from the entrypoint:
   - “Read `dev/instructions/CHAT_INSTRUCTIONS.md` and follow the selected instruction modules.”

## Notes on validation

- `use_instructions()` validates that requested modules exist.
- If you typo a module name, the error will include the list of available modules so you can correct quickly.
